Vagueness and imprecise imitation in signaling games

### 2016-02-15 paper/version_05/paper.tex
Added some motivation for talking about orderly categories.

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell and the young
Wittgenstein which is challenged by the paradoxes vagueness gives rise
to.

@@

This result is interesting because it demonstrates that signaling can
impose orderly categories on a metric space, without that being the
ulterior purpose of it all. Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.

Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.

>>>>

Many concepts and expressions are vague. A vague category knows clear
cases that fall under it, clear cases that do not, and also so-called
borderline cases. Borderline cases do not clearly apply, nor do they
clearly not apply, and there may be differences between borderline
cases in terms of how well they represent the category in
question. Vagueness does not seem to dramatically affect the success
of everyday communication, but it is troublesome for some of the most
prominent theories of language and meaning. This is especially so for the logico-positivist tradition of Frege, Russell, and early
Wittgenstein, which is challenged by the paradoxes vagueness gives rise
to.

@@

It is intuitive to think that linguistic and conceptual categories are orderly in such a manner.
For example, if you consider two people, one with a height of 2m and another with a height of 2.2m, both `tall', it would be difficult to defend considering a person with a height of 2.1m `short'.
Another example is that of color categorization.
The World Color Survey project~\citep{WCS} collected color naming data for 110 unwritten languages of 45 language families.
For the great majority of these languages a pattern can be observed: basic color terms are by and large convex~\citep{Regier07,Jager2010}.
It is premature to argue that these observations can be extended to \emph{all} cases of categorization.
However, for the cases where it does apply, the result of \citet{JagerMetzger2011:Voronoi-Languag} is interesting because it demonstrates that signaling can impose this kind of orderly categories on a metric space without that being the ulterior purpose of it all.

There are at least two potential ways of interpreting the signaling setup.
Sender and receiver can be distinct
entities, whose purpose is to communicate effectively about the actual
state. In that case, evolving Voronoi languages would explain why
\emph{linguistic categories} are well-behaved and orderly in the way
they appear to be. More abstractly, sender and receiver can also be
thought of as distinct modules in a single system, where the first
module must discretize the information it is fed by selecting a small
sample of, suggestively, category labels. These are passed to a second
module that tries to decode the original information. In this case,
evolving Voronoi languages would explain why \emph{conceptual
  categories} are well-behaved and orderly in the way that they appear
to be.
Seen in this light, sim-max games may provide a foundation to approaches in cognitive semantics
that rely on the notion of conceptual spaces.
\citet[][70--77]{Gardenfors2000:Conceptual-Spac}, for example, has prominently argued that
natural categories are convex regions in conceptual space. If the conceptual space has a
suitable metric, convex categories can be derived from a set of prototypes. The category
corresponding to prototype $p$ is the set of points that are more similar to $p$ than to any
other. In this way, \citeauthor{Gardenfors2000:Conceptual-Spac} argues, an efficient categorization
system can be obtained: storing the prototypes lets us recover the categories without having to
store each category's extension. However, what is left unexplained is where the prototypes
come from, and why we would not see just any distribution of prototypes as an equally efficient
classification system. This is where sim-max games can contribute a principled approach to
deriving, in an independent way, not only convex categories but also prototypical exemplars
belonging to them.  These ideas, and more, are developed further by
\citet{Jager2007:The-Evolution-o,JagerRooijvan-Rooij2007:Language-Struct,JagerMetzger2011:Voronoi-Languag,OConnor2014-OCOEPC},
among others.
